---
title: "Exercise 01"
author: "Vebash Naidoo"
date: "29/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    eval = FALSE
)
```

```{r, eval = TRUE}
library(DBI)        # Database interface to connect to a DB
library(tidyverse)  
library(dbplyr)     # Database backend for dplyr
```

## Make a connection

Let's open a connection to the database we have stored locally in our working directory.

---------------- Complete the code ----------------

```{r connection}
con <- dbConnect(
    drv = RSQLite::______,
    dbname = "great_brit_bakeoff.db"
)
```

## Join Data

We're interested in seeing how the winning bakers performed across the season they won. 

- How many times did they win `Star Baker`?
- How many times where they top of the technical charts?

To do that we need to use the tables __baker_results__ and __results__.

### Query the `results` table

Let's get a list of the winners!

```{r}
winners_q <- tbl(con, "results") %>% 
  filter(result == "WINNER") %>% 
  select(series, baker)
```

The above is just a query string, and gives us a glimpse of the data if we execute it, but no actual data resides in R as yet.


### Join the tables

Okay, let's join the two tables and visualise the result.

---------------- Complete the code ----------------

```{r join}
join_q <- tbl(con, "baker_results") %>% 
  _____(winners_q, 
        by = c('baker' = 'baker',
               'series' = 'series'))

join_q %>% 
  show_query() # Have a look at the SQL
```

Now finally we bring in the data into R.

```{r}
winners_data <- 
  # Take our query as-is now
  join_q %>% 
  # add some final adjustments
  select(series, baker:result) %>% 
  # we only want one record per winner
  # since lots are repeated info - so get the
  # record for when they won
  filter(result == "WINNER") %>% 
  # Bring data into R
  collect()
```

### Visualise Data

How did the winner of series 1 do across the season in terms of technical challenges etc.?

```{r}
library(tidyquant)
theme_set(theme_tq())

winners_data %>% 
  # filter for season we're interested in
  filter(series == 1) %>% 
  # let's make all the 
  pivot_longer(cols = c(star_baker:technical_median),
               names_to = "metric") %>% 
  mutate(metric = fct_reorder(metric, value)) %>% 
  ggplot(aes(x = value, y = metric)) +
  geom_col() +
  labs(title = str_glue("Metrics for Season ",
    "{winners_data %>%  filter(series == 1) %>% select(series)}'s Winner ",
    "{winners_data %>%  filter(series == 1) %>% select(baker)}"),
    y = "")
```

### Your Turn

Choose a winner you're interested in looking at.

```{r}
winners_data %>% 
  select(series, baker)
```

Got one? Fill in the blanks below to see how they performed across the season. 

__Note:__ Once you're ready you can amend `eval = FALSE` to be `eval = TRUE` in the code chunk `setup` right at the top of this file, to include all output in your knitted document.

---------------- Complete the code ----------------

```{r your-turn}
winners_data %>% 
  filter(series == ______ ) %>% 
  pivot_longer(cols = c(star_baker:technical_median),
               names_to = "metric") %>% 
  mutate(metric = fct_reorder(metric, value)) %>% 
  ggplot(aes(x = value, y = metric)) +
  geom_col() +
  labs(title = str_glue("Metrics for Season ",
    "{winners_data %>%  filter(series == ______ ) %>% select(series)}'s Winner ",
    "{winners_data %>%  filter(series == ______ ) %>% select(baker)}"),
    y = "")
```

### Disconnect!!

Always remember to disconnect.

```{r}
dbDisconnect(con)
```


