---
title: "Make DB from starwars dataset"
author: "Vebash Naidoo"
date: "29/12/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
library(tidyverse)
```

## Setup a MySQL local server

1. Go to [mysql page](https://dev.mysql.com/downloads/) to download the appropriate server for you. I am on a Windows machine so I used the _MySQL Installer for Windows__.
1. Run through the setup.
1. Follow the setup steps - I installed a "Full version".
1. If you did not setup your "root" user during the setup (this happened to me) open the `MySQL Installer` from your `Startup Menu`.

    a. Click `Reconfigure` next to __MySQL Server__.
    a. Check `Show Advanced and Logging Options`.
    a. Click __Next >__ to navigate the setup, which will include asking you to setup a password for root.

1. I opened `MySQL Workbench` and setup a user named `admin`, after connecting to the local host.
1. I created a __"schema"__ by clicking on the little Database icon in `MySQL Workbench`

    a. Link is "Create a new schema in the connected server". 
    a. Alternately in the `MySQL Command Line Client` run:
    
      > CREATE DATABASE starwars CHARACTER SET utf8 COLLATE utf8_unicode_ci;


That's it for the MySQL Workbench itself.

## Let's create tables from the dplyr::starwars data

Here, I am just pulling apart the different parts of the starwars data to create mini tables instead of one dataframe.

For example, I want a table with all the film names, one with all the vehicle names etc. instead of these as embedded lists.

```{r, eval = TRUE, echo = FALSE}
# data(package = "dplyr")
# starwars %>% 
#   View()

(films <- starwars %>%  
  # get the films column which contains lists of values
  select(films) %>% 
  # flatten the films list - this will create multiple rows for
  # each film in the list
  unnest(films) %>% 
  # get rid of duplicates
  distinct() )

film_years <- tribble(~name,      ~year,
        #-----------#------
        "The Empire Strikes Back", 1980,
        "Revenge of the Sith", 2005,
        "Return of the Jedi", 1983,
        "A New Hope", 1977,
        "The Force Awakens", 2015,
        "Attack of the Clones", 2002,
        "The Phantom Menace", 1999)

films <- films %>% 
  inner_join(film_years,
             by = c("films" = "name"))

film_levels <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith",
                 "A New Hope", "The Empire Strikes Back", "Return of the Jedi",
                 "The Force Awakens")
films <- films %>% 
  mutate(films = factor(films, film_levels),) %>% 
  # create an id column - we will use this later
  # the .before just says I want the id column to come before the films column
  mutate(id = as.integer(films), .before = "films") %>% 
  arrange(id)

(vehicles <- starwars %>%  
  # do the same for vehicles
  select(vehicles) %>% 
  unnest(vehicles) %>% 
  distinct() %>% 
  mutate(id = row_number(), .before = "vehicles"))

(starships <- starwars %>%  
  # do the same for starships
  select(starships) %>% 
  unnest(starships) %>% 
  distinct() %>% 
  # create an id column - we will use this later
  # the .before just says I want the id column to come before the starships column
  # since only 2 columns I will have an id then starships column in this table
  mutate(id = row_number(), .before = "starships"))

appearances <- starwars %>% 
  # let's save the characters appearances in the movies
  select(name, films) %>% 
  # we end up with character_name film for every film the character appears in
  unnest(films) %>%
  # join it with our new films table
  inner_join(films) %>% 
  # select the id which we rename to film_id, and then the name of the character
  select(film_id = id, name)

vehicles_piloted <- starwars %>% 
  select(name, vehicles) %>% 
  unnest(vehicles) %>% 
  inner_join(vehicles) %>% 
  select(vehicle_id = id, name)

starships_piloted <- starwars %>% 
  select(name, starships) %>% 
  unnest(starships) %>% 
  inner_join(starships) %>% 
  select(starship_id = id, name)

characters <- starwars %>% 
  # let's get only the non-list columns
  select(name:species)
```

### 2018 TidyTuesday Star Wars Survey Data

We have some data from a TidyTuesday dataset in [2018 Week 7](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-05-14).

Let's see if we can incorporate some of that data as well.

```{r}
# Read just the first 2 header lines so we can create better column names
starwars_survey_header <- read_csv("week7_starwars.csv", n_max = 2, 
                                col_names = FALSE)

# The headers are quite messy
# Following this blog post entry let's try and clean 'em up
# https://debruine.github.io/posts/multi-row-headers/
starwars_survey_header <- starwars_survey_header %>% 
  # transpose
  t() %>% 
  as_tibble() %>%
  # fill NAs with last non-NA value
  fill(V1, .direction = "down") %>% 
  # combine the V1 and V2 cols
  mutate(name = paste(V1, V2, sep = " ")) %>% 
  pull(name) %>% 
  # get rid of extra whitespace wherever it may be in the string
  str_squish() %>% 
  as_tibble() %>% 
  # the column = value, here we rename to "name"
  rename(name = "value") %>% 
  # many fields have an extra "Response" - we'll remove this
  mutate(name = str_remove_all(name, "Response")) %>% 
  # The main name cleaning - shorten names, prep to match the character names
  # in the starwars dataset
  mutate(name = case_when(
    str_detect(name, "RespondentID NA") ~ 
      "respondent_id",
    str_detect(name, "Have you seen any of the 6 films in the Star Wars franchise\\?") ~ 
      "any_of_6",
    str_detect(name, "Do you consider yourself to be a fan of the Star Wars film franchise\\?") ~
      "star_wars_fan",
    str_detect(name, "Which of the following Star Wars films have you seen\\? Please select all that apply. Star Wars: Episode [IV]+ ") ~ 
      str_replace(name, 
        "Which of the following Star Wars films have you seen\\? Please select all that apply. Star Wars: Episode [IV]+ ", 
        "watched_"),
    str_detect(name, "Please rank the Star Wars films in order of preference with 1 being your favorite film in the franchise and 6 being your least favorite film. Star Wars: Episode [IV]+ ") ~ 
      str_replace(name, 
        "Please rank the Star Wars films in order of preference with 1 being your favorite film in the franchise and 6 being your least favorite film. Star Wars: Episode [IV]+ ", 
        "rank_"),
    str_detect(name, "Please state whether you view the following characters favorably, unfavorably, or are unfamiliar with him/her. ") ~ 
      str_replace(name, 
        "Please state whether you view the following characters favorably, unfavorably, or are unfamiliar with him/her. ", 
        ""),
    str_detect(name, "Which character shot first\\?") ~ 
      "who_shot_first",
    str_detect(name, "Are you familiar with the Expanded Universe?") ~ 
      "know_expanded_universe",
    str_detect(name, "Do you consider yourself to be a fan of the Expanded Universe\\?") ~
      "fan_expanded_universe",
    str_detect(name, "Do you consider yourself to be a fan of the Star Trek franchise\\?") ~
      "trekkie",
    # These don't seem to work, so we will again clean up later
    # I am still unsure what's wrong here so kept in to dissect later
    str_detect(name, "Princess") ~ 
      str_replace(name, "Princess ", ""),
    str_detect(name, "Obi Wan") ~ 
      "Obi-Wan Kenobi",
    str_detect(name, "Emperor") ~ 
      str_replace(name, "Emperor ", ""),
    str_detect(name, "R2 D2") ~ 
      "R2-D2",
    TRUE ~ name
  )) %>% 
  pull(name) %>% 
  str_squish()

# Read the data, skipping the column names
starwars_survey <- read_csv("week7_starwars.csv", skip = 2, 
                                col_names = FALSE)

# reset column names to the cleaned version names
names(starwars_survey) <- starwars_survey_header

# starwars_survey %>% View()

svy_levels <-  c("Very unfavorably", "Somewhat unfavorably", 
                    "Neither favorably nor unfavorably (neutral)",
                    "Somewhat favorably", "Very favorably",
                    "Unfamiliar (N/A)")

survey_levels <- starwars_survey %>% 
  # We're going to create a table for the levels of the survey answers
  select(level = `Han Solo`) %>% 
  distinct() %>% 
  drop_na() %>% 
  mutate(level = factor(level, levels = svy_levels),
         id = as.integer(level)) %>% 
  select(id, level)

character_survey <- starwars_survey %>% 
  select(respondent_id, `Han Solo`:`Yoda`) %>% 
  mutate(across(c(`Han Solo`:`Yoda`), ~factor(., levels = svy_levels))) %>% 
  mutate(across(c(`Han Solo`:`Yoda`), ~as.integer(.)))
  

starwars_survey <- 
  starwars_survey %>% 
  mutate(across(c(starts_with("watched_")), 
                ~if_else(is.na(.), "No", "Yes"))) %>% 
  select(-c(`Han Solo`:`Yoda`)) %>% 
  left_join(character_survey,
            by = c("respondent_id" = "respondent_id")) %>% 
  # fix names
  rename("Leia Organa" = "Princess Leia Organa") %>% 
  rename("Obi-Wan Kenobi" = "Obi Wan Kenobi") %>% 
  rename("Palpatine" = "Emperor Palpatine") %>% 
  rename("R2-D2" = "R2 D2") %>% 
  # reorder cols
  select(respondent_id:`rank_Return of the Jedi`,
        `Han Solo`:`Yoda`,
        everything())
```

We also had some information about the franchise income.

```{r}
media_franchises <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-02/media_franchises.csv")

# Get name of franchise
media_franchises %>% 
  count(franchise, sort = TRUE)

# Filter info for Star Wars
star_wars_franchise_info <- media_franchises %>% 
  filter(franchise == "Star Wars") %>% 
  rename(revenue_billion_dollars = revenue) # keep note that revenue is in $ bil
```


```{r}
library(DBI)
# install.packages("RMariaDB")
library(RMariaDB)

# Make a connection
con <- dbConnect(odbc::odbc(), 
  .connection_string = "Driver={MySQL ODBC 8.0 Unicode Driver};Server=localhost;Port=3306;Database=starwars;Uid=admin;Pwd=Luke1@my0urf@ther;", 
    timeout = 10)

# What tables do we have? Should be none or character(0)
# when you first connect after creating the DB Shell
dbListTables(con)
```

Alternately we can use the driver.

```{r, eval = FALSE}
# Alternate way to make a connection
con <- dbConnect(drv = RMariaDB::MariaDB(), 
                 host = "localhost", 
                 port = "3306",
                 user = "admin", 
                 password = "Luke1@my0urf@ther",
                 dbname = "starwars")
dbListTables(con)
dbDisconnect(con)
```

```{r}
# Let's write all our dataframes to tables
# Using our connection, write to the table "characters"
DBI::dbWriteTable(conn = con, "characters", 
                  # What must be written? - The characters dataframe
                  # If the table exists I want you to wipe its contents
                  # and replace with content from my df characters 
                  characters, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "films",
                  films, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "vehicles",
                  vehicles, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "starships",
                  starships, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "appearances",
                  appearances, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "vehicles_piloted",
                  vehicles_piloted, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "starships_piloted",
                  starships_piloted, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "franchise",
                  star_wars_franchise_info, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "survey_levels",
                  survey_levels, overwrite = TRUE)
DBI::dbWriteTable(conn = con, "survey",
                  starwars_survey, overwrite = TRUE)
```


```{r}
# Okay what tables do we have now?
dbListTables(con)

# Let's confirm we have some data ...
tbl(con, "films")
tbl(con, "survey")

# Good housekeeping - disconnect
dbDisconnect(con)
```


